Annotating, understanding, and predicting long-term video memorability

### 2018-01-26 ARTICLE.tex
Changes in modelling part according to suggestions from Ngoc and Claire-Helene

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MEMORABILITY PREDICTION} %for Khartik
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS AND FUTURE WORKS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%avantages de notre dataset: long-term memorabity
%Inconvénient: mesure avec une part de subjectivité

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MEMORABILITY PREDICTION} %for Khartik
\label{mem-pred}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Until now, we have presented the video collection and the annotation protocol. 
%We then computed memorability scores and analyzed how different users fared in remembering the video sequences from the movies they have already watched.
In this section, we move towards building a machine learning model that can learn and then predict the memorability score of a video from its audio-visual features.
The main goal of modelling is to identify which features: generic, perceptual, or semantic, are important for predicting the memorability scores of a video and \emph{not} the best learner/classifier.
We pose the problem as a standard regression problem and Figure \ref{prop-apprch} illustrates different steps in our method.
First, we see that the videos are converted into feature vectors and then a regression model is trained that can in-turn predict the memorability scores for new videos.
In the following sub-sections, we explain our choice of features and models to address the problem in hand.

\begin{figure}[h]	  
  \centering
    \includegraphics[width=0.9\columnwidth]{figures/approach.pdf}
		\caption{Proposed approach for memorability score prediction.}
    \label{prop-apprch}
\end{figure}

%Before we go into the technical details about the different audio-visual features and regression models we explored, we will talk about how we split our dataset.
To build our predictive model, we split the dataset, at the level of movies, into training (70\%), validation (15\%) and test (15\%) data, which translates into 70 movies in the training set and 15 movies each in the validation and test sets.
We chose to split our dataset at the level of movies, instead of the videos, in order to avoid videos from the same movie being present in the training as well as the evaluation (validation+test) set.

\subsection{Feature extraction}
\label{feat-extr}
%We explore a variety of audio-visual features in building a model that can predict memorability scores for new videos.
The task of remembering a specific sequence from a movie has a high cognitive complexity in general, and is difficult for humans.
Many users who participated in our experiments indicated that it is a difficult task.
While trying to build a machine learning model for such a task, we wanted to explore different kinds of features that could be extracted from the audio-visual signal.
We investigate a variety of generic state-of-the-art features (\cite{c3d-feat}, \cite{audioset-feat}) and compare them with other semantic (\cite{caption-feat}) and perceptual (emotion) features (\cite{sb-feat}).

%We mostly stick to state-of-the-art deep learning features that have been proposed in vision and audio research communities for tasks like video classification \cite{c3d-feat}, event detection \cite{audioset-feat}, image captioning \cite{caption-feat}, and concept detection \cite{sb-feat}.

\subsubsection{Spatio-temporal visual features (C3D)}
\label{c3d-feat}
These features are extracted from the C3D model, a 3-dimensional convolutional network proposed for generic video analysis \cite{c3d-feat}. 
The main motivation to use C3D is that it encodes both the spatial and temporal information in the video.
The model has been proposed for video analysis and is not an extension of a model for image analysis, unlike other state-of-the-art models like VGG16 \cite{vgg16}.
We use the publicly available model trained on the Sports-1M dataset \cite{c3d-feat}.
We use the output of the fully connected layer -- fc6 of the network with a dimensionality of 4096.
Since such dimensionality is high compared to other features, we explore the use of principal component analysis (PCA) for the dimension reduction.

\subsubsection{Audio features (AudioSet)}
\label{as-feat}
Using a recently released large-scale dataset of manually annotated events: AudioSet \cite{audioset-feat} and the corresponding event detection models, we extract 128-dimensional embeddings for each audio track associated with a video in our dataset.
We use these embeddings for training the regression models.
The motivation to use these features is that they are state-of-the-art in the audio event detection research and events could play a major role in how people remember sequences in movies.

\subsubsection{Emotion related features (SentiBank and Affect)}
\label{emo-feat}
As research in psychology showed that emotion and memory are highly correlated \cite{emo-mem}, we investigate the use of emotion-related feature in our prediction system.
For emotion from visual content, we resort to a visual sentiment concept detector: SentiBank \cite{sb-feat}.
SentiBank is a set of 1200 trained visual concept detectors providing a mid-level representation of sentiment from visual content.
We use the binary code for concept detection, from images, provided by the authors.
We sample one frame for every second of the video in our dataset, resulting in 10 frames per video.
For each of these 10 frames, we run the SentiBank concept detector and obtain two pieces of information: concepts with probabilities and features.
Concepts are adjective--noun pairs and the probability represents how likely this concept is depicted visually in that particular frame.
Examples of some concepts in the SentiBank ontology are: young driver, scary face, terrible pain, etc.
We rank the concepts based on the probability of their occurrence in the frame and take the top-50 concepts.
We extract 300-dimensional word2vec \cite{word2vec} embeddings, for each of the 50 concepts and take an average to obtain a single vector per frame.
We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
Sentibank detectors also provide 4096-dimensional features for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
In the end, we use two features: 300-dimensional concept vectors and 4096-dimensional feature vectors.

Emotion will also be embedded in the audio signal and hence we resort to an audio-visual analysis of the video to obtain arousal and valence scores \cite{affect}.
Arousal is the dimension of emotion that measures the excitement in the video, while valence measures whether the video invokes positive or negative emotion, following a circumplex model of affect \cite{affect-model}.
For each frame in the video, we compute the arousal and valence scores using the method proposed in \cite{affect}.
In order to keep a fixed dimensionality of the feature vector, we take the first 200 frames in the video because of the varying frame rates across the videos.
We concatenate the arousal and valence scores for the first 200 frames in each video resulting in a 400-dimensional feature vector (200 for arousal and 200 for valence) for a video.

\subsubsection{Visual semantic features (Image captions)}
\label{sem-feat}
Visual semantics are known to play an important role in memorability \cite{sem-mem}.
We utilize the state-of-the-art research in image captioning to capture the semantics \cite{caption-feat}.
We sample one frame for every second of the video in our dataset, resulting in 10 frames per video.
For each of these 10 frames, we run the caption detector (code provided by the authors) and obtain a caption for the frame.
For each non-functional word in the caption, we extract 300-dimensional word2vec \cite{word2vec} embeddings and taking an average of all the words to obtain a single vector per frame.
We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single 300-dimensional feature vector for each video.

\subsection{Modelling and evaluation}
\label{model-eval}
We used the features discussed in Section \ref{feat-extr} to train an SVR model with two different kernels (linear and RBF) for the memorability score prediction. 
We use the grid search strategy to obtain the best hyper-parameters for SVR: $C$ = ${0.1, 1, 10, 100, 1000}$ and $\gamma$ = ${0.01, 0.1, 1, 10, 100}$.
The choice of SVR is guided by the small size of the dataset and we stick to the two standard kernels because our focus was mainly on identifying which features are more important for memorability prediction.
We have chosen to go with the same classifier for all the features so as to ensure that the differences in performances is because of the feature themselves.

Since we are posing the memorability prediction problem as a regression problem, we use the standard regression metric: Mean squared error ($MSE$) for evaluation.
Additionally, we also use the spearmann correlation ($SpCorr$) to measure the rank correlation between the predicted memorability scores and the ground-truth.
This metric gives us an indication of the correlation between the predicted memorability scores and the ground-truth memorability scores.

\subsection{Memorability prediction results}
\label{pred-res}
%In this section, we will discuss how the models trained on different features perform in predicting memorability scores of new videos.
In addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating all the feature vectors into a single feature vector.
While concatenating the feature vectors, we use the lower dimensional features for C3D and SentiBank features, obtained after applying PCA to the original set of features.
%Each of the videos in our dataset has multiple annotations and we pick sequences with at least 4 annotations for training the models.
%In this way, we eliminate movies that were not seen by a majority of the users who participated in our experiment.
%Since we are using an SVR with linear and rbf kernels, we train the models on the training set (70 movies) and validate the model on the validation set (15 movies), before reporting the final scores on the test set (15 movies).
We pick the SVR model parameters: kernel (linear or rbf), $C$ and $\gamma$ that give the best performance in terms of $MSE$.
We retrain the model with these parameters on the training set and evaluate on the final test set.
In the experiments where we use a dimension reduction method (PCA) for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.

We report the results of the prediction on both the validation and test sets, so as to investigate whether there is overfitting of the model.
We also report only correlation scores: $SpCorr$ as there is not much change in the $MSE$ scores across different features.
On an average, the $MSE$ score varies between 0.05 and 0.08 across different features.
Table \ref{res-4-10-ann} reports the the results of prediction capability of the model trained on different features.
There are two sets of results reported in the table: $SpCorr$ (4 annotations) and $SpCorr$ (8 annotations).
The first set of results correspond to the prediction capability of the model when trained on videos with at least 4 annotations and the second set corresponds to the results when trained on videos with at least 8 annotations.
We will further discuss why we report the results with the model trained on videos with at least 8 annotations.
Observing the first set of results in Table \ref{res-4-10-ann}, we can clearly see that the image captions and C3D occupy the top-2 places in terms of performance on the test set. 
Image captions capture the visual semantics in the video, while C3D features encode the visual spatio-temporal information.
These results are consistent with previous results on predicting image memorability \cite{squalli_2017_deep}.
Our dataset consists of videos taken from movies that have a specific story-line; semantics and spatio-temporal information seem to be playing an important role in people remembering specific scenes from movies.
Other features like AudioSet come close to the performance of the above features (third best feature), but only audio signal does not seem to be enough for predicting video memorability.
Another observation from Table \ref{res-4-10-ann} is that the combination of all the features (last row in the table) does not appear in the top-3 best performing features.
One of the reasons for this could be that there is a lot of redundance when combining all the features into a single feature vector.
In future, we could look at selectively combining the features to investigate if that improves the performance.

One of our initial hypotheses was that emotion would play a important role in memorability of a video, supported by literature from psychology.
We used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
Observing the scores in Table \ref{res-4-10-ann}, we can say that the SentiBank features (fourth row from bottom) perform slightly better than the affect features (sixth row from bottom).
But the models trained on image caption features clearly out-perform those trained on emotion features.
This could be because of the following two reasons: our choice of features to capture emotion related information is not suitable for our task, or the performance of the emotion prediction models is not good enough for memorability prediction.
From the analysis on this dataset and the features we used for capturing emotion related information, we \emph{cannot} conclude that emotion plays an important role in memorability.
In the future, we could investigate more robust emotion/affect prediction models to improve the performance of memorability prediction.

We further investigate the effect of number of annotations, for videos in the training set, on the performance of the model.
Based on our experimental results, we observe that the image caption feature performs the best while using training videos that have at least 4 annotations.
Based on this result, we train different SVR models with varying number of annotations on the training set using image caption features.
For example, we train a model, using image caption features, on training sequences with at least 4 annotations and use this model to predict the memorability score for videos in the validation set.
We repeat this process for different number of annotations: from 4 to 15 annotations increasing the number of annotations by 1 every time.
Please note that the validation set in each of the repetitions is fixed and only the training set changes.
We observed that the $MSE$ does not change significantly in each of the steps and hence we provide a demonstration of how $SpCorr$ varies with an increasing number of annotations in the training set.
Figure \ref{num-ann} demonstrates the result of this experiment, where we observe that beyond 8 annotations the value of $SpCorr$ keeps going down.
In the wake of this observation, we investigate the performance of all the features when we train the regression model with videos that have at least 8 annotations.

\begin{figure}[h]	  
  \centering
    \includegraphics[width=0.7\columnwidth]{figures/annotations.pdf}
		\caption{Spearmann correlation on the validation set with models trained using image caption features for varying number of annotations.}
    \label{num-ann}
\end{figure}

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    \textbf{Feature} & \textbf{Feature type}& \textbf{Dimension}& \multicolumn{2}{c|}{\textbf{$SpCorr$(4 annotations)}} & \multicolumn{2}{c|}{\textbf{$SpCorr$ (8 annotations)}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{4-7}
    & & & \emph{validation set}& \emph{test set} & \emph{validation set}& \emph{test set}\\ \hline
    C3D& visual-spatio-temporal& 4096& 0.20& 0.26& 0.31& 0.34\\ \hline
		C3D (PCA)& visual-spatio-temporal (PCA)& 225& 0.24& 0.21& 0.18& 0.17\\ \hline
		AudioSet& audio-related& 128& 0.23& 0.22& 0.21& 0.24\\ \hline
		Affect& affect-related& 400& 0.19& 0.17& 0.26& 0.23\\ \hline
		SentiBank concepts& emotion-related& 300& 0.16& 0.13& 0.15& 0.17\\ \hline
		SentiBank features& emotion-related& 4096& 0.25& 0.21& 0.27& 0.26\\ \hline
		SentiBank features (PCA)& emotion-related (PCA)& 225& 0.22& 0.21& 0.22& 0.23\\ \hline
		Captions& visual-semantics & 300& 0.29& 0.31& 0.39& 0.38 \\ \hline
		Combination& combine-all-features (PCA)& 1578& 0.24& 0.23& 0.29& 0.27\\ \hline		
    \hline
  \end{tabular}
  \caption{Prediction results (Spearmann correlation) on validation and test data for different features with models trained on sequences that have at least 4 (columns 4-5) or 10 (columns 6-7) annotations.}
	\label{res-4-10-ann}	
\end{table*}

These results are reported in the second part of Table \ref{res-4-10-ann}: $SpCorr$ (8 annotations).
Comparing the two sets of results in Table \ref{res-4-10-ann}, we observe that the model trained on sequences with at least 8 annotations performs better than the model trained on sequences with at least 4 annotations.
The only exception being the C3D (PCA) feature, where the model trained on sequences with at least 4 annotations performs marginally better than the model trained on sequences with at least 8 annotations.

Additionally, we also observe that the scores for validation set and test set are very similar across the two results and hence we can say that there is no overfitting.
We also compute the average number of annotations per sequence in the training and test set for both the cases (4 and 8 annotations) in order to check if there is a mismatch.
On an average, each sequence in the test set has around 10 annotations and the test set remains the same across different experiments.
Similarly, each sequence in the training set has 9 annotations on an average in the case where we train the models with sequences having at least 4 annotations and 11 annotations in the other case.
Looking at these numbers, we can see that the average number of annotations per sequence is quite balanced across the training and test sets and there is indeed no mismatch.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS AND FUTURE WORKS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%avantages de notre dataset: long-term memorabity
%Inconvénient: mesure avec une part de subjectivité
After exploring a variety of generic (C3D, AudioSet), perceptual (SentiBank, Affect), and semantic (Image captions) features, we can say that a model trained with  semantic features provides predictions that are most correlated with ground-truth memorability scores.
At this moment, we cannot claim that this would hold true for any other type of videos because we have only investigated one specific type of videos i.e., movies.
Future research could involve investigating the memorability prediction performance on other generic videos using the existing set of features as well as other important features.
We also found that emotion features are not that well correlated with ground-truth memorability scores, as indicated in psychology literature.
An immediate direction for future research would be to explore other approaches to encode emotion in videos and investigate whether we can improve the prediction performance using these approaches.
We also investigated the number of annotations required to train a model and found a sweet spot where the performance is optimal.
